---
title: "BDA400 Assignment 2 - Technical Analysis using R (Preliminary Stage)"
author: "Marwah Mahsal Khan"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Setup

All the functions live in `stock_functions.R` so this file just loads them
and runs them. `quantmod` and `TTR` get loaded inside that script.

```{r load-functions}
source("stock_functions.R")
```

## Portfolio

The stock symbols are stored in `portfolio.txt`, one per line.

```{r portfolio}
symbols <- read_portfolio("portfolio.txt")
symbols
```

## Importing the Data

`load_stock_data()` reads `portfolio.txt` and downloads daily prices for each
symbol from Yahoo Finance (Jan 1 - Dec 31, 2025). Each stock ends up as its
own data frame inside a named list.

```{r load-data}
stocks <- load_stock_data("portfolio.txt", from = "2025-01-01", to = "2026-01-01")

names(stocks)
sapply(stocks, class)   # each one is a separate data frame
sapply(stocks, nrow)    # number of trading days per stock
```

Structure of one of the data frames:

```{r structure}
str(stocks[[1]])
```

The data is also saved as CSV files in an `output/` folder.

```{r save-data}
save_stock_data(stocks)
```

## Displaying the Data

### Historical data (last 10 trading days)

Same layout as the Historical Data tab on Yahoo Finance, newest day first.

```{r historical, results='asis'}
for (sym in names(stocks)) {
  cat("\n\n####", sym, "\n\n")
  print(knitr::kable(display_historical(stocks[[sym]], sym, n = 10),
                     align = "r"))
}
```

### Quote summary

A summary box like the one at the top of a Yahoo Finance stock page. "Period
Range" is the lowest low and highest high over the whole year.

```{r quote-summary, results='asis'}
for (sym in names(stocks)) {
  cat("\n\n####", sym, "\n\n")
  print(knitr::kable(display_quote_summary(stocks[[sym]], sym)))
}
```

## Statistics

`calculate_statistics()` works on the daily closing price for each stock:
mean, median, mode, standard deviation, min, max, and the latest 20-day and
50-day simple moving averages. `calculate_all_statistics()` runs it on every
stock and puts the results into one table.

```{r statistics}
stats <- calculate_all_statistics(stocks)
knitr::kable(stats)
```

A few notes on how these are calculated:

- **Mode** - base R doesn't have a mode function, and closing prices basically
  never repeat to the exact cent, so prices are rounded to the nearest dollar
  first and the most common value is taken.
- **SD** - `sd()` gives the sample standard deviation (divides by n - 1).
- **Moving averages** - `SMA()` from the TTR package. The value shown is from
  the last trading day of the year.

## Visualizations

### Closing price with moving averages

```{r price-ma, fig.width=8, fig.height=4.5}
for (sym in names(stocks)) {
  plot_price_with_ma(stocks[[sym]], sym)
}
```

### Candlestick charts with volume (quantmod)

```{r candlestick, fig.width=8, fig.height=5}
for (sym in names(stocks)) {
  plot_candlestick(stocks[[sym]], sym)
}
```

### Relative performance of the whole portfolio

Every stock starts at 100 on the first trading day so they can be compared on
one chart.

```{r normalized, fig.width=8, fig.height=4.5}
plot_normalized(stocks)
```

## Session Info

```{r session-info}
sessionInfo()
```
